Flowchart of Current Algorithm

### 2008-08-15 docs/flowchart/paper.tex
merged the thermal diffusion stuff in

In paper II we discretized the density and enthalpy evolution equations in 
perturbational form, arguing that this would be less
susceptible to grid effects.  In the present algorithm, the analog would
be to use a perturbational form of the species equations.
Numerical testing has shown that the form of the algorithm presented
here is in fact more robust than the perturbational form 
when the base state mixing terms ($\etarho$) are included.  
In the algorithm described below, density is still 
extrapolated to edges in perturbational form, but 
we treat $X_k$ and $h$ in non-perturbational form.

@@

%--------------------------------------------------------------------------
% STEP 2
%--------------------------------------------------------------------------
\noindent {\bf Step 2.} {\em Construct the provisional edge-based advective velocity}, $\uadvone$:

@@

\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2),\star}$ and
$(\rho h)^{(1)} \rightarrow (\rho h)^{(2),\star}$ without explicitly including the 
reaction terms,
\begin{eqnarray}
(\rho X_k)^{(2),\star} &=& (\rho X_k)^{(1)} 
 - \dt \; \left[ \nablab \cdotb \left( (\uadvone+w_0^{\nph,\star} \er)  
  (\rho X_k)^{(1),\nph,\star} \right) \right] \enskip , \\ \nonumber  \\
(\rho h)^{(2),\star} &=& (\rho h)^{(1)} - \dt \; \left[ \nablab \cdotb \left( (\uadvone+w_0^{\nph,\star} \er)  
  (\rho h)^{(1),\nph,\star} \right) \right] \nonumber \\
&& + \dt \; \left(\uadvone \cdotb \er\right) \left(\frac{\partial p_0}{\partial r} \right)^{n} 
+ \dt \; \psi^{\nmh} \enskip , \\
\rho^{(2),\star} &=& \sum_k (\rho X_k)^{(2),\star} \enskip , \\ 
X_k^{(2),\star} &=& (\rho X_k)^{(2),\star} / \rho^{(2),\star} \enskip , \\
T^{(2),\star} &=& T\left(\rho^{(2),\star}, (\rho h)^{(2),\star},  X_k^{(2),\star}\right) {\hbox{ \rm using the equation of state.}} 
\end{eqnarray}

@@

@@

The procedure to construct $\uadvtwo$ is described in detail in Appendix B of paper III
and is analogous to the procedure used in step 2, but with updated values
for $w_0$ and $\pizero.$ We note that  $\uadvtwo$ satisfies the discrete versions of 
equations (\ref{eq:udoterzero}) and (\ref{eq:tildeconstraint}),
in particular
\begin{equation}
\nablab \cdotb \left(\beta_0^{\nph,\star} \uadvtwo\right) = \beta_0^{\nph,\star} 
\left(S^{\nph} - \overline{S}^{\nph}\right) \enskip .
\end{equation}
where $ \beta_0^{\nph,\star} = ( \beta_0^n +  \beta_0^{n+1,\star}) / 2.$

@@

\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2)}$ and
$(\rho h)^{(1)} \rightarrow (\rho h)^{(2)},$ 
\begin{eqnarray}
(\rho X_k)^{(2)} &=& (\rho X_k)^{(1)} 
- \dt \; \left[ \nablab \cdotb \left( (\uadvtwo+w_0^{\nph} \er)  
(\rho X_k)^{(1),\nph} \right) \right]  \enskip , \\ \nonumber \\
(\rho h)^{(2)} &=& (\rho h)^{(1)} - \dt \; \left[ \nablab \cdotb \left( (\uadvtwo+w_0^{\nph} \er)  
(\rho h)^{(1),\nph} \right) \right] \nonumber \\
&& + \frac{\dt}{2} \; \left(\uadvtwo \cdotb \er\right)
\left [ \left(\frac{\partial p_0}{\partial r} \right)^{n}
      + \left(\frac{\partial p_0}{\partial r} \right)^{n+1}  \right ] 
+ \dt \; \psi^{\nph,\star} \enskip , \\ 
\rho^{(2)} &=& \sum_k (\rho X_k)^{(2)} \enskip , \\
X_k^{(2)} &=& (\rho X_k)^{(2)} / \rho^{(2)} \enskip , \\
T^{(2)} &=& T\left(\rho^{(2)}, (\rho h)^{(2)},  X_k^{(2)}\right) {\hbox{ \rm using the
equation of state.}} 
\end{eqnarray}

@@

\item Define an edge-centered $ \etarho^{\nph}$ and a cell-centered $\psi_j^{\nph}$,
\begin{eqnarray}
 \etarho^{\nph}   &=&  {\rm {\bf Avg}} \left( (\uadvtwo \cdotb \er + w_0^{\nph}) 
\rho^{(1),\nph} \right) - w_0^{\nph} \rho_0^{\nph,{\rm pred}} 
 \nonumber \enskip , \\
 \psi_j^{\nph}  &=&  \frac{1}{2} \left(\eta_{rho,j-\myhalf}^{\nph} + \eta_{\rho,j+\myhalf}^{\nph}\right) g \enskip .
\end{eqnarray}
\end{enumerate}

@@

The initial perturbational pressure also needs to be determined for
use in steps 2, 7 and 11. 
This is done through a second iterative procedure which follows the
time advancement algorithm as described in steps 1-11 in \S~ref{Sec:Alg}.  
The user specifies the number of iterations, 
$N_{\rm iters}^{\pi},$ in this second step of the initialization procedure.
The details for both iterations are given below.

%--------------------------------------------------------------------------
% STEP 0
%--------------------------------------------------------------------------
\noindent {\bf Step 0} {\em Initialization}

>>>>

In paper II we discretized the density and enthalpy evolution equations in 
perturbational form, arguing that this would be less
susceptible to grid effects.  In the present algorithm, the analog would
be to use a perturbational form of the species equations.
Numerical testing has shown that the form of the algorithm presented
here is in fact more robust than the perturbational form 
when the base state mixing terms ($\etarho$) are included.  
In the algorithm described below, density is still 
extrapolated to edges in perturbational form, but 
we treat $X_k$ and $h$ in non-perturbational form.\\

@@

%--------------------------------------------------------------------------
% STEP 2
%--------------------------------------------------------------------------
\noindent {\bf Step 2.} {\em Construct the provisional edge-based advective velocity}, $\uadvone$.\\

@@

\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2),\star}$ and
$(\rho h)^{(1)} \rightarrow (\rho h)^{(2),\star}$ without explicitly including the 
reaction terms,
\begin{eqnarray}
(\rho X_k)^{(2),\star} &=& (\rho X_k)^{(1)} 
 - \dt \; \left[ \nablab \cdotb \left( (\uadvone+w_0^{\nph,\star} \er)  
  (\rho X_k)^{(1),\nph,\star} \right) \right] \enskip , \\ \nonumber  \\
(\rho h)^{(2'),\star} &=& (\rho h)^{(1)} - \dt \; \left[ \nablab \cdotb \left( (\uadvone+w_0^{\nph,\star} \er)  
  (\rho h)^{(1),\nph,\star} \right) \right] \nonumber \\
&& + \dt \; \left(\uadvone \cdotb \er\right) \left(\frac{\partial p_0}{\partial r} \right)^{n} 
+ \dt \; \psi^{\nmh} \enskip , \\
\rho^{(2'),\star} &=& \sum_k (\rho X_k)^{(2),\star} \enskip , \\ 
X_k^{(2),\star} &=& (\rho X_k)^{(2),\star} / \rho^{(2),\star}
\end{eqnarray}

@@

%--------------------------------------------------------------------------
% STEP 4.1
%--------------------------------------------------------------------------

\item Diffuse the enthalpy through a time interval of $\dt$.  Compute $\kappa^{(1)}, c_p^{(1)}$, and $\xi_k^{(1)}$ from $\rho^{(1)}, T^{(1)}$, and $X_k^{(1)}$ as inputs to the equation of state.  Note that density is constant in this step, i.e., $\rho^{(2'),\star}=\rho^{(2),\star}$.  The update is given by
\begin{eqnarray}
(\rho h)^{(2),\star} &=& (\rho h)^{(2'),\star} + \frac{\dt}{2}\nabla\cdot\left(\frac{\kappa^{(1)}}{c_p^{(1)}}\nabla h^{(2),\star} + \frac{\kappa^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\right)\nonumber\\
&&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla X_k^{(2),\star} + \frac{\xi_k^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
&&- \frac{\dt}{2}\nabla\cdot\left(\frac{h_p^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla p_0^{(2),\star} + \frac{h_p^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla p_0^{(1)}\right),
\end{eqnarray}
which is numerically implemented as a diffusion equation for $h^{(2),\star}$,
\begin{eqnarray}
\left(\rho^{(2),\star} - \frac{\dt}{2}\nabla\cdot\frac{\kappa^{(1)}}{c_p^{(1)}}\nabla\right)h^{(2),\star} &=& (\rho h)^{(2'),\star} + \frac{\dt}{2}\nabla\cdot\frac{\kappa^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\nonumber\\
&&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla X_k^{(2),\star} + \frac{\xi_k^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
&&- \frac{\dt}{2}\nabla\cdot\left(\frac{h_p^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla p_0^{(2),\star} + \frac{h_p^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla p_0^{(1)}\right),
\end{eqnarray}
\item Update the temperature
\begin{equation}
T^{(2),\star} = T\left(\rho^{(2),\star}, h^{(2),\star}, X_k^{(2),\star}\right) {\hbox{ \rm using the equation of state.}}
\end{equation}

@@

The procedure to construct $\uadvtwo$ is described in detail in Appendix B of paper III
and is analogous to the procedure used in step 2, but with updated values
for $w_0$ and $\pizero.$ We note that  $\uadvtwo$ satisfies the discrete versions of 
equations (\ref{eq:udoterzero}) and (\ref{eq:tildeconstraint}),
in particular
\begin{equation}
\nablab \cdotb \left(\beta_0^{\nph,\star} \uadvtwo\right) = \beta_0^{\nph,\star} 
\left(S^{\nph} - \overline{S}^{\nph}\right) \enskip .
\end{equation}
where $ \beta_0^{\nph,\star} = ( \beta_0^n +  \beta_0^{n+1,\star}) / 2.$\\

@@

\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2)}$ and
$(\rho h)^{(1)} \rightarrow (\rho h)^{(2)},$ 
\begin{eqnarray}
(\rho X_k)^{(2)} &=& (\rho X_k)^{(1)} 
- \dt \; \left[ \nablab \cdotb \left( (\uadvtwo+w_0^{\nph} \er)  
(\rho X_k)^{(1),\nph} \right) \right]  \enskip , \\ \nonumber \\
(\rho h)^{(2')} &=& (\rho h)^{(1)} - \dt \; \left[ \nablab \cdotb \left( (\uadvtwo+w_0^{\nph} \er)  
(\rho h)^{(1),\nph} \right) \right] \nonumber \\
&& + \frac{\dt}{2} \; \left(\uadvtwo \cdotb \er\right)
\left [ \left(\frac{\partial p_0}{\partial r} \right)^{n}
      + \left(\frac{\partial p_0}{\partial r} \right)^{n+1}  \right ] 
+ \dt \; \psi^{\nph,\star} \enskip , \\ 
\rho^{(2')} &=& \sum_k (\rho X_k)^{(2)} \enskip , \\
X_k^{(2)} &=& (\rho X_k)^{(2)} / \rho^{(2)}
\end{eqnarray}

@@

\item Define an edge-centered $ \etarho^{\nph}$ and a cell-centered $\psi_j^{\nph}$,
\begin{eqnarray}
 \etarho^{\nph}   &=&  {\rm {\bf Avg}} \left( (\uadvtwo \cdotb \er + w_0^{\nph}) 
\rho^{(1),\nph} \right) - w_0^{\nph} \rho_0^{\nph,{\rm pred}} 
 \nonumber \enskip , \\
 \psi_j^{\nph}  &=&  \frac{1}{2} \left(\eta_{rho,j-\myhalf}^{\nph} + \eta_{\rho,j+\myhalf}^{\nph}\right) g \enskip .
\end{eqnarray}

%--------------------------------------------------------------------------
% STEP 8.1
%--------------------------------------------------------------------------
\item Diffuse the enthalpy through a time interval of $\dt$.  Compute $\kappa^{(2),\star}, c_p^{(2),\star}$, and $\xi_k^{(2),\star}$, from $\rho^{(2),\star}, T^{(2),\star}$, and $X_k^{(2),\star}$ as inputs to the equation of state.  Note that density is constant in the diffusion step, i.e., $\rho^{(2')}=\rho^{(2)}$.  The update is given by
\begin{eqnarray}
(\rho h)^{(2)} &=& (\rho h)^{(2')} + \frac{\dt}{2}\nabla\cdot\left(\frac{\kappa^{(2),\star}}{c_p^{(2),\star}}\nabla h^{(2)} + \frac{\kappa^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\right)\nonumber\\
&&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(2),\star}\kappa^{(2),\star}}{c_p^{(2),\star}}\nabla X_k^{(2)} + \frac{\xi_k^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
&&- \frac{\dt}{2}\nabla\cdot\left(\frac{h_p^{(2),\star}\kappa^{(2),\star}}{c_p^{(2),\star}}\nabla p_0^{(2)} + \frac{h_p^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla p_0^{(1)}\right),
\end{eqnarray}
which is numerically implemented as a diffusion equation for $h^{(2)}$,
\begin{eqnarray}
\left(\rho^{(2)} - \frac{\dt}{2}\nabla\cdot\frac{\kappa^{(2),\star}}{c_p^{(2),\star}}\nabla\right)h^{(2)} &=& (\rho h)^{(2')} + \frac{\dt}{2}\nabla\cdot\frac{\kappa^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\nonumber\\
&&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(2),\star}\kappa^{(2),\star}}{c_p^{(2),\star}}\nabla X_k^{(2)} + \frac{\xi_k^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
&&- \frac{\dt}{2}\nabla\cdot\left(\frac{h_p^{(2),\star}\kappa^{(2),\star}}{c_p^{(2),\star}}\nabla p_0^{(2)} + \frac{h_p^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla p_0^{(1)}\right),
\end{eqnarray}
\item Update the temperature
\begin{equation}
T^{(2)} = T\left(\rho^{(2)}, h^{(2)}, X_k^{(2)}\right) {\hbox{ \rm using the equation of state.}}
\end{equation}

@@

The initial perturbational pressure also needs to be determined for
use in steps 2, 7 and 11. 
This is done through a second iterative procedure which follows the
time advancement algorithm as described in steps 1-11 in \S~ref{Sec:Alg}.  
The user specifies the number of iterations, 
$N_{\rm iters}^{\pi},$ in this second step of the initialization procedure.
The details for both iterations are given below.\\

%--------------------------------------------------------------------------
% STEP 0
%--------------------------------------------------------------------------
\noindent {\bf Step 0.} {\em Initialization}
